---
title: "Practical machine learning"
author: "Mike Vosskuhler"
date: "September 5, 2018"
output: html_document
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(caret)
library(randomForest)
library(rpart)
library(gbm)
```

# This project  
The aim of this project is to demonstrate some basic machine learning methods,
used to predict outcomes of a small test set. 
To train the model we will use a much larger training set.  

## downloading and loading data  
The first step is to download and load some data  
```{r cache=T}
# (down)loading data
dir <- "C:/Users/vossk/Practical_machineLearning"
setwd(dir)
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
              destfile = "training.csv")
download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",
              destfile = "testing.csv")
training <- read.csv("training.csv")
testing <- read.csv("testing.csv")
```

# Exploration and cleaning  
Next we will explore and clean the data.  
Our exploratory analysis showed that the data contained lots of NAs and empty values.  
We will remove the columns containing these missing values.   
```{r cache=T}
str(training) #dataset contains lots of NAs
# removing any data containing NAs
NAFRAC <- function(x){sum(is.na(x)| x == "")/length(x)}
v <- (apply(training, 2, NAFRAC))
train_subs <- training[,!v]
train_subs <- train_subs[, -(1:7)]
```

## Data partition  
Next we create a data partition for training and testing.   
```{r cache=T}
set.seed(1234)
intrain <- createDataPartition(train_subs$classe, p = 3/4, list = F)
Train <- train_subs[intrain,]
Test <- train_subs[-intrain,]
```

## Model building  
Because of limited computational resources we use only use one method:  

### Rpart
Here we build a model using the rpart algorithm, then we test the accuracy. 

```{r cache=T}
model1 <- train(classe~., data = Train, method = "rpart", na.action = na.pass)
pred <- predict(model1, Test)
confusionMatrix(pred, Test$classe)
```
the accuracy here is only approaximately 50 percent.
This is not great and would not be acceptable in a real live experiment, however,
for now we will accept this model and predict the testing data using model1.  

```{r}
finalpred <- predict(model1, testing)
print(finalpred)
```